Random Forests", BUAs

### 2024-02-08 src/bua_RF.Rmd
bua

---
title: "Random Forests", OAs
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

@@

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## spatial data OLD

```{r eval=F}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
colnames(oa.ew@data) <- c('id', 'area')

# reassign the polygon IDs
oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.ew)

# # get OA for Scotland
# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
# 
# # spatial transformations
# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
# 
# # Scotland (follows same steps as EW, see notes above)
# oa.sc <- oa.sc[, 'DataZone']
# colnames(oa.sc@data) <- c('id')
# 
# # reassign the polygon IDs
# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
# 
# # check the CRS has changed correctly, and the data slot has shrink to only the ID
# summary(oa.sc)

path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
colnames(oa.sc@data) <- c('id', 'area')

# reassign the polygon IDs
oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.sc)

# build oa for GB
oa.gb <- maptools::spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

# oa.gb$geometry <- oa.gb$geometry %>%
#   s2::s2_rebuild() %>%
#   sf::st_as_sfc()

# get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- readOGR(path.gb)
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))
```

## spatial data NEW

```{r}
library(sf)

@@

# NI
path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
  rename(id= SA2011,
         area = Hectares)

# Build OA for UK
oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
rm(oa.ew, oa.sc, oa.ni)

# Get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- st_read(path.gb)

# Spatial transformations
gb <- st_transform(gb, 4326)  # EPSG code for WGS84
```

@@

oa.full <- replicate(17, oa.uk %>% as_tibble(), simplify = FALSE) %>%
  bind_rows(.id = 'id_') %>% 
  rename(year = id_) %>% 
  mutate(year = as.numeric(year) + 1995) %>% 
  dplyr::select(id, year)

@@

## Distances to cities and retail centres

### Cities as points

First set up the destinations. Origins are all the OA.

Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)

I use the centroids from arcgis API as they better reflect the city centres than 
than the geometric centres of the boundaries, which I use for `sum(n)`

```{r}
# oa centroids
# oa.uk.c <- (rgeos::gCentroid(oa.uk, byid=TRUE))@coords %>% as.data.frame()
# oa.uk.c$id <- row.names(oa.uk.c)  #as_tibble(rownames = NA) %>% 
#   mutate(id = row.names())

# oa.uk.c <- gCentroid(oa.uk, byid=TRUE)@coords %>% 
#   as.data.frame() %>% 
#   add_rownames(var = "id")
sf_use_s2(FALSE)
oa.uk.c <- st_centroid(oa.uk) %>%
  st_coordinates() %>%
  as.data.frame() %>%
  #add_rownames(var = "id") %>% 
  bind_cols(oa.uk %>% dplyr::select(id))

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK",
"Belfast, UK")

cities <- geo(city.names, no_query = F, method = "arcgis") %>% 
  mutate(address = stringr::str_remove(address, ", UK"))

# # Replace London with Inner London centroid
# # Source: https://data.london.gov.uk/dataset/sub-regions-london-plan-consultation-2009
# london.path <- paste0(path, "/data/raw/London_shape")
# london.sf <- st_read(dsn = london.path, layer = "lp-consultation-oct-2009-subregions")
# # Spatial transformations
# london.sf <- st_transform(london.sf, 4326)  # EPSG code for WGS84
# 
# london.c <- london.sf %>% 
#   filter(Name == "Central") %>% 
#   st_centroid() %>% 
#   st_coordinates() %>%
#   as.data.frame() %>% 
#   rename(lat = Y, long = X) %>% 
#   mutate(address = "London, UK") %>% 
#   dplyr::select(address, lat, long)
# 
# cities <- cities %>% bind_rows(london.c) %>% 
#   slice(2:n()) %>% # to remove the old London centroid based on the cities geometry
#   mutate(address = stringr::str_remove(address, ", UK"))
```

### City boundaries

@@

# # cities
# city.names <- c(
#   "London", 
#   "Birmingham",
#   "Glasgow",
#   "Liverpool",
#   "Bristol",
#   "Manchester",
#   "Sheffield",
#   "Leeds",
#   "Edinburgh",
#   "Leicester",
#   )

# City boundaries
# Source: https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=name&tags=all(BDY_TCITY%2CDEC_2015)
path.in <- paste0(path, "/data/raw/Major_Towns_and_Cities_Dec_2015_Boundaries_V2_2022_121402779136482658.geojson")

city.names <- city.names %>% stringr::str_remove(", UK")

city.boundaries.ew <- st_read(path.in) %>% 
  st_transform(4326) %>% 
  filter(TCITY15NM %in% city.names) %>% 
  dplyr::select(OBJECTID, TCITY15CD, TCITY15NM) %>% 
  rename(code = TCITY15CD,
         name = TCITY15NM)

# Source: https://geoportal.statistics.gov.uk/
path.in <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
city.boundaries.sc.ni <- st_read(path.in) %>% 
  st_transform(4326) %>% 
  filter(grepl('Edinburgh|Glasgow|Belfast', LAD21NM)) %>% 
  dplyr::select(OBJECTID, LAD21CD, LAD21NM) %>% 
  rename(code = LAD21CD,
         name = LAD21NM)

# merge
city.boundaries <- rbind(city.boundaries.ew, city.boundaries.sc.ni)
```

### Retail centres

```{r}
# retail centres <- 
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
#retail <-readOGR(geo.path) 
retail <- st_read(geo.path) %>% st_transform(4326)  # EPSG code for WGS84

#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
#   as.data.frame() %>% 
#   add_rownames(var = "id") #%>%
retail.major.cetres.help <- st_centroid(retail) %>% 
  #st_transform(4326) %>%
  st_coordinates() %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# retail.major.cetres <- retail.major.cetres.help %>% 
#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
#   st_transform(4326) %>%
#   st_coordinates() %>%
#   as_tibble() %>% 
#   bind_cols(retail.major.cetres.help$id) %>% 
#   rename(id = '...3') %>% 
#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
retail.major.cetres <- retail.major.cetres.help %>% 
  # as_tibble() %>% 
  # bind_cols(retail.major.cetres.help$id) %>% 
  # #rename(id = '...3') %>% 
  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
```

@@

```{r}
# NI
oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))

dist.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist.ni <- round((dist.ni/1000),2) %>% 
  as_tibble()

names(dist.ni) <- cities$address #city.names

dist.ni <- dist.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

@@

# min distance
dist.ni <- transform(dist.ni, dist = do.call(pmin, dist.ni[-1]))

# City name for minimum distance 
dist.ni <- dist.ni %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])

# GB
oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))

dist.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist.gb <- round((dist.gb/1000),2) %>% 
  as_tibble()

names(dist.gb) <- cities$address #city.names

dist.gb <- dist.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# min distance
dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-1]))

# City name for minimum distance 
dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])

# UK
dist <- rbind(dist.ni, dist.gb)

# calculate area
#oa.uk$area <- st_area(oa.uk) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
  rename(oa11cd=id)

@@

### Distance to retail centres

```{r}
# dist.retail <- distm(cbind(oa.uk.c$X, oa.uk.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
# dist.retail <- round((dist.retail/1000),2) %>% 
#   as_tibble()

# NI
oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
dist.retail.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail.ni <- round((dist.retail.ni/1000),2) %>% 
  as_tibble()

names(dist.retail.ni) <- retail.major.cetres$RC_ID

dist.retail.ni <- dist.retail.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# This efficiently returns the name of the column with the shortest distance
dist.retail.ni <- dist.retail.ni %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])

# dist.retail_ <- dist.retail %>% slice_sample(n=100)
# dist.retail_[,6425:6426]
# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
# sum(is.na(dist.retail_$dist.retail.name))

# Minimum distance 
dist.retail.ni <- transform(dist.retail.ni, dist.retail = do.call(pmin, dist.retail.ni[2:6424])) 
dist.retail.ni <- dist.retail.ni  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)

# GB
oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
dist.retail.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail.gb <- round((dist.retail.gb/1000),2) %>% 
  as_tibble()

names(dist.retail.gb) <- retail.major.cetres$RC_ID

dist.retail.gb <- dist.retail.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# This efficiently returns the name of the column with the shortest distance
dist.retail.gb <- dist.retail.gb %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])

# dist.retail_ <- dist.retail %>% slice_sample(n=100)
# dist.retail_[,6425:6426]
# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
# sum(is.na(dist.retail_$dist.retail.name))

# Minimum distance 
dist.retail.gb <- transform(dist.retail.gb, dist.retail = do.call(pmin, dist.retail.gb[2:6424])) 
dist.retail.gb <- dist.retail.gb  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)

# UK
dist.retail <- rbind(dist.retail.ni, dist.retail.gb)

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd")) 
#sapply(df, function(x) sum(is.na(x)))
```

## n for London, nearest city and retail

### TODELETE

```{r, eval=FALSE}
# ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
# #df <- df %>% filter(!is.na(area)) 
# 
# cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
# cities.sf <- st_set_crs(cities.sf, 4326)
# 
# # help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
# #   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
# 
# sf_use_s2(FALSE)
# help1 <- st_filter( oa.gb, cities.sf) %>% 
#   #st_coordinates() %>% 
#   as_tibble() %>% 
#   dplyr::select(id) %>% 
#   mutate(cities = c("London",     # The city names have been manually checked and assigned
#                     "Leicester",
#                     "Leeds",
#                     "Sheffield",
#                     "Bristol",
#                     "Birmingham",
#                     "Manchester",
#                     "Liverpool",
#                     "Edinburgh",
#                     "Glasgow"))
# 
# # too slow
# #dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
# 
# # This efficiently returns the name of the column with the shortest distance
# dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
# 
# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
#   #rename(nearest.city.oa11cd = id) %>% 
#   dplyr::select(oa11cd, dist.city.name)
# 
# df <- df %>%  left_join(help2, by = c("oa11cd" = "oa11cd")) 
```

### n for cities

```{R}
#help.retail <- dist.retail

# n nearest city
# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
#                          rename(n.nearest.city = n), 
#                        by = c("nearest.city.oa11cd" = "oa11cd",
#                               "year" = "year"))

sf_use_s2(FALSE)
city.boundaries.oa <- st_join(oa.uk, city.boundaries, join = st_intersects) %>% 
  as_tibble() %>% 
  filter(!is.na(name)) %>% 
  dplyr::select(id, name) %>% 
  mutate(name = ifelse(name == "City of Edinburgh", "Edinburgh", name)) %>% 
  mutate(name = ifelse(name == "Glasgow City", "Glasgow", name))

city.boundaries.help <- df %>% inner_join(city.boundaries.oa, by = c("oa11cd" = "id")) %>% 
  group_by(name, year) %>% 
  summarise(n.nearest.city = sum(n)) %>% 
  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
  ungroup()

london.boundaries.help <- city.boundaries.help %>% 
  filter(name == "London") %>% 
  rename(n.London = n.nearest.city,
         n.London.lag = n.nearest.city.lag)

df <- df %>% left_join(city.boundaries.help, by = c("dist.city.name" = "name", "year" = "year"))

df <- df %>% left_join(london.boundaries.help, by = c("year" = "year"))

#sapply(df, function(x) sum(is.na(x)))
```

### DELETE n for London

```{r, eval=FALSE}

# # OLD London definitions
# # df <- df %>% left_join(df %>% 
# #                          ungroup() %>% 
# #                          filter(oa11cd == "E00176659") %>% 
# #                          dplyr::select(year, n) %>% 
# #                          rename(n.London = n)) %>% 
# #   group_by(oa11cd) %>% 
# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
# #   mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
# #   ungroup() 
# # 
# # london.help <- df %>% group_by(RGN11NM, year) %>% 
# #   summarise(n.London = sum(n)) %>% 
# #   filter(RGN11NM == "London") %>%
# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
# #   ungroup() %>% 
# #   dplyr::select(-RGN11NM) 
# 
# london.oa <- st_filter(oa.gb, london.sf %>% 
#                          filter(Name == "Central")) %>% 
#   as_tibble() %>% 
#   dplyr::select(id)
# 
# london.help <- df %>% inner_join(london.oa, by = c("oa11cd" = "id")) %>% 
#   group_by(year) %>% 
#   summarise(n.London = sum(n)) %>% 
#   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
#   ungroup()
# 
# df <- df %>% left_join(london.help, by = c("year" = "year"))
```

### n for nearest retail

```{r}
# Old definition for retail centre
# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
#                          rename(n.nearest.retail = n), 
#                        by = c("oa11cd.retail" = "oa11cd",
#                               "year" = "year")) %>% 
#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
# 
# 
# #n.nearest.retail IS WRONG. FIX UPSTREAM
# df <- df %>% 
#   rename(oa11cd.retail = n.nearest.retail) %>% 
#   dplyr::select(-n.nearest.retail.lag) %>% 
#   left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
#                          rename(n.nearest.retail = n), 
#                        by = c("oa11cd.retail" = "oa11cd",
#                               "year" = "year")) %>% 
#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
# 
# 
# df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
#   left_join(london.help, by = c('year' = 'year'))

retail.oa <- st_join(oa.uk, retail, 
                     join = st_intersects) %>% #, 
                    #largest = TRUE) %>% # returns the x features augmented with 
                                         # the fields of y that have the largest 
                                         # overlap with each of the features of x
  as_tibble() %>% 
  filter(!is.na(RC_ID)) %>% 
  dplyr::select(id, RC_ID) %>% 
  add_row(id = "E00019779", RC_ID = "RC_EW_5951")

# If `larget = FALSE`, we have multiple RC_IDs for OAs.
# Also, RC_ID="RC_EW_5951" would fall in Thames as it is the London Bridge Tube Station.
# To avoid NAs for for n.nearest.retail, I would have assigned dist.retail.name = RC_EW_1935 for E00019779

# st_join(oa.uk, retail, join = st_is_within_distance, dist = 500) %>% 
#   as_tibble() %>% 
#   filter(!is.na(RC_ID)) %>% 
#   dplyr::select(id, RC_ID)
# 
# st_filter(oa.uk, retail,  .pred = st_is_within_distance, dist = 100) %>% 
#   as_tibble() %>% 
#   filter(!is.na(RC_ID)) %>% 
#   dplyr::select(id, RC_ID)

# h <- oa.uk %>% st_intersects(retail)

# retail.oa_ %>% filter(id == "E00004828"|id == "E00004829"|id == "E00004827")
# # RC_EW_5285
# retail.oa <- retail.oa %>% add_row(id = "E00004828", RC_ID = "RC_EW_5285")
# retail.oa <- retail.oa %>% arrange(id)

retail.help <- df %>% full_join(retail.oa, by = c("oa11cd" = "id")) %>% 
  group_by(RC_ID, year) %>% 
  summarise(n.nearest.retail = sum(n)) %>% 
  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail)) %>% 
  ungroup()

# retail.help2 <- retail.oa %>% left_join(retail.help, by = c("RC_ID" = "RC_ID")) %>% 
#   dplyr::select(-RC_ID)

df <- df %>% left_join(retail.help, by = c("dist.retail.name" = "RC_ID", "year" = "year"))
#sapply(df, function(x) sum(is.na(x)))
```

### Urban Rural TODO

## TODO

```{r}
# Lookup: OA to regions
# **FIXED< DELETE IN TEST** NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION

# # regional identifier
# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
#   ungroup() %>% 
#   dplyr::select(-help)

#sapply(df, function(x) sum(is.na(x)))
```

@@

```{r}
# sf_use_s2(FALSE)
# oa.c <- st_centroid(oa.uk)
# kn <- knearneigh(oa.c, k = 5)
# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
# knn.l <- nb2listw(knn)

path.out <- paste0(path, "/data/temp/knnl.RData")
# save(knn.l, file = path.out)
load(path.out)

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
```

@@

@@

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #%>% #growth, abs.growth, RGN11NM, 
  # mutate(London.v = n.London.lag / London,
  #        city.v = n.nearest.city.lag / dist,
  #        retail.v = n.nearest.retail.lag / (dist.retail + .1)) %>% 
  # dplyr::select(-n.London.lag, -London, -n.nearest.city.lag, -dist, -n.nearest.retail.lag, -dist.retail)

@@

# df %>% dplyr::select(area, dist, dist.retail, London, south, n.slag, n.lag, n.l.slag, n.nearest.city, year) %>% 
#   #sapply(function(x) sum(is.na(x)))
#   summarise(across(where(is.numeric), .fns = 
#                      list(#n = is.na,
#                           min = min,
#                           median = median,
#                           mean = mean,
#                           stdev = sd,
#                           # q25 = ~quantile(., 0.25),
#                           # q75 = ~quantile(., 0.75),
#                           max = max))) %>%
#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
# 
# df %>% dplyr::select(n, n.London.lag, n.nearest.city.lag, n.nearest.retail.lag, n.l.slag, year, London, dist, dist.retail) %>%
#   #sapply(function(x) sum(is.na(x)))
#   summarise(across(where(is.numeric), .fns =
#                      list(#n = is.na,
#                           min = min,
#                           median = median,
#                           mean = mean,
#                           stdev = sd,
#                           # q25 = ~quantile(., 0.25),
#                           # q75 = ~quantile(., 0.75),
#                           max = max))) %>%
#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
# 
# df %>% slice_max(order_by = n, n = 5, by = year)
```

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "oa11cd",
                                timevar = "year", 
                                k = 10)

@@

@@

### OLD

```{r, eval=FALSE}
# train on all but one region, test on that region

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)

>>>>

---
title: "Random Forests", BUAs
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

TODO:

- check missing bua
- check out
- regions
- growth

@@

# temp for working on laptop
# path.in <- paste0(path, "/data/temp/help.oa.csv")
# df <- read_csv(path.in)

# https://geoportal.statistics.gov.uk/datasets/df834d6d654d4f4a9a211fdbbb8893bd_0/explore
# path.lookup.rg <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")
# lookup.rg <- read_csv(path.lookup.rg) %>% 
#   dplyr::select(-ObjectId)
```

## spatial data

@@

# Build OA for GB
oa.gb <- rbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

# # Get GB
# path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
# gb <- st_read(path.gb)
# 
# # Spatial transformations
# gb <- st_transform(gb, 4326)  # EPSG code for WGS84
```

@@

oa.full <- replicate(17, oa.gb %>% as_tibble(), simplify = FALSE) %>%
  bind_rows(.id = 'id_') %>% 
  rename(year = id_) %>% 
  mutate(year = as.numeric(year) + 1995) %>% 
  dplyr::select(id, year)

@@

## OA to BUA

```{r}
# long to wide
df.w <- df %>% pivot_wider(names_from = year, values_from = n)

sf_use_s2(FALSE)
oa.gb.c <- st_centroid(oa.gb) %>%
  left_join(df.w)

# BUA
# source: https://geoportal.statistics.gov.uk/datasets/ad30b234308f4b02b4bb9b0f4766f7bb_0/explore
bua.path <- paste0(path, "/data/raw/BUA_2022_GB_-8042218937152150708.geojson")
bua <- st_read(bua.path)
bua <- st_transform(bua, 4326)  # EPSG code for WGS84

# OA fall within a BUA
df.within <- st_join(oa.gb.c, bua, join = st_intersects) %>% #, join = st_intersects
  as_tibble()

df.out <- df.within %>% filter(is.na(OBJECTID))
df.within <- df.within %>% filter(!is.na(OBJECTID))

# help <- st_is_within_distance(oa.gb.c %>% filter(id %in% df.out$id), bua, dist=500) %>% #, join = st_intersects
#   as_tibble()
df.out <- st_join(oa.gb.c %>% filter(id %in% df.out$id), bua, join = st_nearest_feature) %>% #, join = st_intersects
  as_tibble()

df <- bind_rows(df.within, df.out) %>% # 142953
  dplyr::select(BUA22CD, BUA22NM, 3:19) %>% 
  pivot_longer(!c(BUA22CD, BUA22NM), names_to = "year", values_to = "n") %>% 
  group_by(BUA22CD, year) %>% 
  summarise(n = sum(n)) %>% 
  ungroup() %>% 
  left_join(bua %>% as_tibble() %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
  relocate(BUA22NM, .after = BUA22CD) %>% 
  mutate(year = as.numeric(year))

# # WE ARE MISSING 42855 websites were in OA not in a BUA
# # The above might help but takes a long time
# df <- st_join(oa.gb.c, bua, join = st_intersects) %>% # 94265
#   as_tibble() %>% 
#   dplyr::select(BUA22CD, BUA22NM, 3:19) %>% 
#   pivot_longer(!c(BUA22CD, BUA22NM), names_to = "year", values_to = "n") %>% 
#   group_by(BUA22CD, year) %>% 
#   summarise(n = sum(n)) %>% 
#   ungroup() %>% 
#   left_join(bua %>% as_tibble() %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
#   relocate(BUA22NM, .after = BUA22CD) %>% 
#   mutate(year = as.numeric(year))

sapply(df, function(x) sum(is.na(x)))  
# df <- df %>% filter(is.na(BUA22CD))
# 42855 websites were in OA not in a BUA (3,536,635 - 3,493,780)
```

## Distances to cities

### City BUA

@@

city.names <- c("City of Westminster",
"Birmingham",
"Glasgow",
"Liverpool",
"Bristol",
"Manchester",
"Sheffield",
"Leeds",
"Edinburgh",
"Leicester")

#cities.bua <- df %>% filter(BUA22NM %in% city.names) %>% distinct(BUA22CD)
cities.bua <- bua %>% filter(BUA22NM %in% city.names)

cities.bua.c <- cities.bua %>% filter(BUA22NM %in% city.names) %>% 
  st_centroid() %>%
  st_coordinates() %>%
  as.data.frame() %>%
  #add_rownames(var = "id") %>% 
  bind_cols(cities.bua %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
  dplyr::select(BUA22CD, BUA22NM, X, Y)

bua.c <- bua %>% 
  st_centroid() %>%
  st_coordinates() %>%
  as.data.frame() %>%
  #add_rownames(var = "id") %>% 
  bind_cols(bua %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
  dplyr::select(BUA22CD, BUA22NM, X, Y)
```

@@

dist.gb <- distm(cbind(bua.c$X, bua.c$Y), cbind(cities.bua.c$X, cities.bua.c$Y), fun=distHaversine) 
dist.gb <- round((dist.gb/1000),2) %>% 
  as_tibble()

names(dist.gb) <- cities.bua.c$BUA22NM #city.names

dist.gb <- dist.gb %>% 
  mutate(BUA22CD = bua.c$BUA22CD,
         BUA22NM = bua.c$BUA22NM) %>% 
  relocate(c(BUA22CD, BUA22NM))

@@

# min distance
dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-(1:2)]))

# City name for minimum distance 
# +2L to accommodate for the first two columns: BUA22CD and BUA22NM
dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[3:12]*-1)+2L])

# Join with distance and area
df <- df %>% left_join(dist.gb %>% dplyr::select(-BUA22NM))

@@

## n for London and nearest city

```{R}
df <- df %>% mutate(dist.city.name = ifelse(dist.city.name == "City.of.Westminster",
                                             "City of Westminster", dist.city.name)) %>% 
  left_join(cities.bua.c %>%
              dplyr::select(BUA22CD, BUA22NM) %>% 
              rename(BUA22CD.city = BUA22CD),
            by = c("dist.city.name" = "BUA22NM")) %>% 
  left_join(df %>% dplyr::select(BUA22CD, year, n) %>% 
              rename(n.nearest.city = n),
            by = c("BUA22CD.city" = "BUA22CD", "year" = "year")) %>% 
  left_join(df %>% dplyr::select(BUA22NM, year, n) %>% 
              rename(n.London = n) %>% 
              filter(BUA22NM == "City of Westminster") %>% 
              rename(help = BUA22NM),
            by = c("year" = "year")) %>% 
  dplyr::select(-c(help, Leeds, Liverpool, Manchester, Sheffield, Leicester, Birmingham, 
                   Bristol, Edinburgh, Glasgow))
```

@@

bua.used <- df %>% distinct(BUA22CD) %>% as.vector()
sf_use_s2(FALSE)
bua.c <- bua %>% filter(BUA22CD %in% bua.used$BUA22CD) %>% 
  st_centroid()
kn <- knearneigh(bua.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(BUA22CD) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.London.lag = dplyr::lag(n.London, n=1, order_by = year),
         n.nearest.city.lag = dplyr::lag(n.nearest.city, n=1, order_by = year))
```

@@

NOT DONE

@@

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(BUA22CD, BUA22NM, year, n, n.l.slag, n.nearest.city.lag, 
                n.London.lag, City.of.Westminster, dist,) %>% 
  rename(London = City.of.Westminster)

@@

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "BUA22CD",
                                timevar = "year", 
                                k = 10)

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # It should be RHS variables + 2

@@
